Characterisation of transient resonances in extreme-mass-ratio-inspirals

### 2014-08-06 asymptotic.tex
more small changes

The following derivation is based upon the analysis of Kevorkian~\cite{Kevorkian1987}.\footnote{The same two time-scale theory underpins the analysis of Hinderer and Flanagan~\cite{Hinderer2008}, but this explicitly ignores resonances.} Small adjustments have been made to adapt to the specific problem of GW inspiral, but the general argument is unchanged.

We model the system using action--angle variables. We are only concerned with the $r$ and $\theta$ motions, so we have a $2$-dimensional system. We perform a canonical transformation  to isolate the resonant combination $q = n_r q_r - n_\theta q_\theta$~\cite{Bosley1992}. This becomes one of the new angle variables, the other variable $q'$ can be either $q_r$ or $q_\theta$ (as, on resonance, varying one necessarily changes the other). We use $J$ as the conjugate action variable to $q$ and $\omega = n_r \omega_r - n_\theta \omega_\theta$ as its frequency. Similarly, we use $J'$ as the action variable conjugate to $q'$. The system evolves through resonance slowly, on an evolution time-scale, so we parameterize it in terms of a slow time parameter
\begin{equation}
\widetilde{\lambda} = \eta\lambda.
\end{equation}
The orbits of $q'$ proceed with the fast time $\lambda$; since this is much more rapid than the evolution we are interested in, it is safe to average over it. We are not interested in the fine-grained fast oscillations caused by changes in $q'$. For this analysis we consider the reduced problem of evolving $q$ and $J$.

@@

To find a solution away from the resonance, we decompose the problem to be a function of two time-scales~\cite{Kevorkian1971}. We use the slow time $\widetilde{\lambda}$ and, as a proxy for the fast time,
\begin{equation}
\Psi = \intd{0}{\lambda}{\omega(\eta\tau)}{\tau} = \recip{\eta}\intd{0}{\tilde{\lambda}}{\omega(\widetilde{\tau})}{\widetilde{\tau}}.
\end{equation}
From this
\begin{equation}
\omega = \diff{\Psi}{\lambda}.
\end{equation}
In terms of these two variables, we can build ansatz solutions
\begin{subequations}
\begin{align}
\label{eq:q-series}
q(\lambda;\,\eta) = {} & \Psi + q_0\left(\widetilde{\lambda}\right) + \eta q_1\left(\Psi,\widetilde{\lambda}\right) + \order{\eta^2}, \\
J(\lambda;\,\eta) = {} & J_0\left(\widetilde{\lambda}\right) + \eta J_1\left(\Psi,\widetilde{\lambda}\right) + \order{\eta^2}.
\label{eq:J-series}
\end{align}
\end{subequations}
We can also write a series expansion for the frequency, since it is affected by the self-force too,
\begin{equation}
\omega(\lambda;\,\eta) = \omega_0\left(\widetilde{\lambda}\right) + \eta \omega_1\left(\widetilde{\lambda}\right) + \order{\eta^2}.
\end{equation}
In the limit of $\eta \rightarrow 0$ we are left with a constant frequency $\omega_0(0)$. The higher order terms are identified below by matching terms in the series expansions of the equations of motion. Taking the two time-scales as independent, we may write the time derivative to $\order{\eta}$ as
\begin{equation}
\diffop{\lambda} = \omega_0\partialdiffop{\Psi} + \eta\omega_1\partialdiffop{\Psi} + \eta\partialdiffop{\widetilde{\lambda}}.
\end{equation}

Using the two time-scale decomposition to replace the time derivatives in the equations of motion, and substituting in the ansatz expansions gives, to first order,
\begin{widetext}
\begin{subequations}
\begin{align}
\label{eq:q-1}
\omega_0 + \eta\omega_1 + \eta\partialdiff{q_0}{\widetilde{\lambda}} + \eta\omega_0\partialdiff{q_1}{\Psi} = {} & \omega(J_0) + \eta\diff{\omega}{J}J_1 + \eta \sum_s g_s^{(1)}(J_0)\exp\left[is(\Psi + q_0)\right], \\
\eta\partialdiff{J_0}{\widetilde{\lambda}} + \eta\omega_0\partialdiff{J_1}{\Psi} = {} & \eta \sum_s G_s^{(1)}(J_0)\exp\left[is(\Psi + q_0)\right].
\label{eq:J-1}
\end{align}
\end{subequations}
\end{widetext}
Averaging \eqnref{J-1} over $\Psi$ gives\footnote{The ansatz is constructed such that $J_0 \equiv \langle J_0\rangle_\Psi$ and $q_0 \equiv \langle q_0\rangle_\Psi$.}
\begin{equation}
\partialdiff{J_0}{\widetilde{\lambda}} = G_0^{(1)}(J_0).
\label{eq:J-ad}
\end{equation}
This describes the adiabatic evolution, hence we can identify $J_0\left(\widetilde{\lambda}\right)$ with (the lowest order piece of) the adiabatic solution~\cite{Hinderer2008}. If we similarly average \eqnref{q-1} we find
\begin{equation}
\omega_0 + \eta\omega_1 +\eta\partialdiff{q_0}{\widetilde{\lambda}} = \omega(J_0) + \eta\partialdiff{\omega}{J}\langle J_1\rangle_\Psi + \eta g_0^{(1)}(J_0).
\end{equation}
From this we can identify the terms that originate from the frequency and, matching by order in $\eta$, obtain
\begin{subequations}
\begin{align}
\omega_0 = {} & \omega(J_0), \\
\omega_1 = {} & \partialdiff{\omega}{J}\langle J_1\rangle_\Psi.
\end{align}
\end{subequations}
This leaves
\begin{align}
\partialdiff{q_0}{\widetilde{\lambda}} = {} & g_0^{(1)}(J_0), \\
q_0 = {} & \kappa_0 + \intd{0}{\tilde{\lambda}}{g_0^{(1)}[J_0(\tau)]}{\tau}.
\label{eq:q-0-sol}
\end{align}
We now have expressions for the lowest order terms in the expansions.

>>>>

The following derivation is reproduced from Berry~\cite{BerryThesis2013} and is based upon the analysis of Kevorkian~\cite{Kevorkian1987}.\footnote{The same two time-scale theory underpins the analysis of Hinderer and Flanagan~\cite{Hinderer2008}, but this explicitly ignores resonances.} Small adjustments have been made to adapt to the specific problem of GW inspiral, but the general argument is unchanged. A similar derivation can be found by van de Meent~\cite{VanDeMeent2014}.

We model the system using action--angle variables. We are only concerned with the $r$ and $\theta$ motions, so we have a $2$-dimensional system. We perform a canonical transformation  to isolate the resonant combination $q = n_r q_r - n_\theta q_\theta$~\cite{Bosley1992,VanDeMeent2014}. This becomes one of the new angle variables, the other variable $q'$ can be either $q_r$ or $q_\theta$ (as, on resonance, varying one necessarily changes the other). We use $J$ as the conjugate action variable to $q$ and $\omega = n_r \omega_r - n_\theta \omega_\theta$ as its frequency. Similarly, we use $J'$ as the action variable conjugate to $q'$. The system evolves through resonance slowly, on an evolution time-scale, so we parameterize it in terms of a slow time parameter
\begin{equation}
\widetilde{\lambda} = \eta\lambda.
\end{equation}
The orbits of $q'$ proceed with the fast time $\lambda$; since this is much more rapid than the evolution we are interested in, it is safe to average over it. We are not interested in the fine-grained fast oscillations caused by changes in $q'$. For this analysis we consider the reduced problem of evolving $q$ and $J$.

@@

To find a solution away from the resonance, we decompose the problem to be a function of two time-scales~\cite{Kevorkian1971}. We use the slow time $\widetilde{\lambda}$ and, as a proxy for the fast time,
\begin{equation}
\Psi = \intd{0}{\lambda}{\omega(\eta\tau)}{\tau} = \recip{\eta}\intd{0}{\tilde{\lambda}}{\omega(\widetilde{\tau})}{\widetilde{\tau}}.
\end{equation}
From this
\begin{equation}
\omega = \diff{\Psi}{\lambda}.
\end{equation}
In terms of these two variables, we can build ansatz solutions
\begin{subequations}
\begin{align}
\label{eq:q-series}
q(\lambda;\,\eta) = {} & \Psi + q_0\left(\widetilde{\lambda}\right) + \eta q_1\left(\Psi,\widetilde{\lambda}\right) + \order{\eta^2}, \\
J(\lambda;\,\eta) = {} & J_0\left(\widetilde{\lambda}\right) + \eta J_1\left(\Psi,\widetilde{\lambda}\right) + \order{\eta^2}.
\label{eq:J-series}
\end{align}
\end{subequations}
We can also write a series expansion for the frequency, since it is affected by the self-force too,
\begin{equation}
\omega(\lambda;\,\eta) = \omega_0\left(\widetilde{\lambda}\right) + \eta \omega_1\left(\widetilde{\lambda}\right) + \order{\eta^2}.
\end{equation}
In the limit of $\eta \rightarrow 0$ we are left with a constant frequency $\omega_0(0)$. The higher-order terms are identified below by matching terms in the series expansions of the equations of motion. Taking the two time-scales as independent, we may write the time derivative to $\order{\eta}$ as
\begin{equation}
\diffop{\lambda} = \omega_0\partialdiffop{\Psi} + \eta\omega_1\partialdiffop{\Psi} + \eta\partialdiffop{\widetilde{\lambda}}.
\end{equation}

Using the two time-scale decomposition to replace the time derivatives in the equations of motion, and substituting in the ansatz expansions gives, to first order,
\begin{widetext}
\begin{subequations}
\begin{align}
\label{eq:q-1}
\omega_0 + \eta\omega_1 + \eta\partialdiff{q_0}{\widetilde{\lambda}} + \eta\omega_0\partialdiff{q_1}{\Psi} = {} & \omega(J_0) + \eta\diff{\omega}{J}J_1 + \eta \sum_s g_s^{(1)}(J_0)\exp\left[is(\Psi + q_0)\right], \\
\eta\partialdiff{J_0}{\widetilde{\lambda}} + \eta\omega_0\partialdiff{J_1}{\Psi} = {} & \eta \sum_s G_s^{(1)}(J_0)\exp\left[is(\Psi + q_0)\right].
\label{eq:J-1}
\end{align}
\end{subequations}
\end{widetext}
Averaging \eqnref{J-1} over $\Psi$ gives\footnote{The ansatz is constructed such that $J_0 \equiv \langle J_0\rangle_\Psi$ and $q_0 \equiv \langle q_0\rangle_\Psi$.}
\begin{equation}
\partialdiff{J_0}{\widetilde{\lambda}} = G_0^{(1)}(J_0).
\label{eq:J-ad}
\end{equation}
This describes the adiabatic evolution, hence we can identify $J_0\left(\widetilde{\lambda}\right)$ with (the lowest order piece of) the adiabatic solution~\cite{Hinderer2008}. If we similarly average \eqnref{q-1}, we find
\begin{equation}
\omega_0 + \eta\omega_1 +\eta\partialdiff{q_0}{\widetilde{\lambda}} = \omega(J_0) + \eta\partialdiff{\omega}{J}\langle J_1\rangle_\Psi + \eta g_0^{(1)}(J_0).
\end{equation}
From this we can identify the terms that originate from the frequency and, matching by order in $\eta$, obtain
\begin{subequations}
\begin{align}
\omega_0 = {} & \omega(J_0), \\
\omega_1 = {} & \partialdiff{\omega}{J}\langle J_1\rangle_\Psi.
\end{align}
\end{subequations}
This leaves
\begin{align}
\partialdiff{q_0}{\widetilde{\lambda}} = {} & g_0^{(1)}(J_0), \\
q_0 = {} & \kappa_0 + \intd{0}{\tilde{\lambda}}{g_0^{(1)}[J_0(\tau)]}{\tau}.
\label{eq:q-0-sol}
\end{align}
We now have expressions for the lowest order terms in the expansions.

### 2014-08-06 res_char.tex
more small changes

To investigate how resonances would impact real-world observations of EMRIs using GWs, we need to generate a realistic astrophysical population of EMRIs\footnote{The actual distribution in parameter space of observable EMRI systems is unknown, but we can make reasonable estimates.} and then evolve these systems in time, including a mechanism to generate the GWs at infinity, which can then be analysed.

@@

\begin{figure}[htbp]
\centering
\includegraphics[width=0.46\textwidth]{Fig_overlap_vs_time}
\caption{\label{fig:overlap-dephasing}The overlap computed between the full evolution and an adiabatic evolution for an illustrative EMRI system with $p=7.85$, as a function of time, including only the parts of the waveforms within some $\Delta t$ of $t$, chosen to give 25 independent (non-overlapping) calculations. The time axis is centred on the value at which the full evolution crosses the 2:3 resonance.}
\end{figure}

@@

To be able to quantify the effect that the resonance has on an evolution, we turn to the SNR and the overlap between full and adiabatic models computed for the entire length of the observed inspiral. The dephasing means that an adiabatic evolution with the same initial conditions as a full evolution is likely to have a low overlap if a significant resonance is encountered. However, this does not necessarily mean that all adiabatic evolutions will have low overlaps; we need to search for optimal adiabatic parameters, such that the overlap is maximal.

@@

The most restrictive aspect of this approach is that all extrinsic parameters and any intrinsic ones that remain constant throughout the evolution, such as the masses of the BHs and the central BH spin, do not get explored by our family of adiabatic models. It is possible that small deviations from these true parameters will give rise to better matching adiabatic models, to partially correct for the (lack of) resonance effects. Indeed, systematic biases are expected when the model waveforms do not accurately reflect the data~\cite{Cutler2007}.

The problem of searching over adiabatic templates now reduces to the task of choosing appropriate values of $t_{\mathrm{match}}$. Anticipating that resonances will play a key role in determining the degree of waveform overlap, we choose matching times situated at $5$ resonance-timescales after each resonance of interest, namely the low-order 2:3, 1:2, 2:5 and 1:3. These matching times lie in a portion of the evolution that is not affected by a resonance and so should result in a large overlap for that region of the inspiral. We also include templates that match at the start of the evolution, so that we have a model that should match the portion of the evolution before the first resonance, and at the end of the evolution, as this is often where most SNR is accumulated. In addition, we include models that match exactly on each of the resonances as these may achieve partial matching both before and after the resonance and hence perform well.

@@

Given our computed resonance jumps for all of the orbital parameters, we extract a single phase parameter $q$ for each system that specifies the magnitude of each jump. We expect such a phase parameter to be linearly offset from $\widehat{\kappa}_0$. The resonance jumps are sinusoidal with respect to $q$, as shown by the fits in \figref{resjump-vs-q}, lending support to \eqnref{delta-I-a} and giving us confidence in our numerical jump calculations. We can also compute the jumps in the fundamental frequencies for each system (shown as dashed lines in \figref{resjump-vs-q}); they are highly correlated to the magnitude of the jump in $E$, due to the dependence on $u_t$.

@@

Strong resonances can severely limit our ability to recover SNR from waveforms using adiabatic templates; they partition the inspiral, splitting up the total SNR into different adiabatic modes, which may be individually undetectable. To assess the impact of this on future gravitational wave missions, we need to create a sample population of detectable EMRIs.

We first identify 14 parameter distributions that describe an astrophysically typical EMRI. We choose the mass of the compact object $\mu = 10 M_\odot$, corresponding to a typical mass for stellar BHs. We take the central object to be a typical supermassive BH at the centre of a galaxy, with mass $M$ uniformly distributed in log-space between $10^5$ and $10^7 M_\odot$, and dimensionless spin $a$ uniformly between its limiting values, $0$ and $1$. The direction of the spin of the central BH, the direction of the angular momentum of the orbit and the direction of the source on the sky are all chosen uniformly on the sphere. The angles specifying the location of pericenter and the compact object around the orbit are distributed uniformly between $0$ and $2\pi$.

@@

These EMRIs are astrophysically motivated, but may not be detectable by eLISA. To assess this, we evolve the systems forward in time using the analytic kludge method of Barack and Cutler \cite{Barack2004} until they reach the last stable orbit. We then calculate the Peters-Matthews waveform \cite{Peters1963} for the last $t_\mathrm{insp}$ years of the inspiral, where $t_\mathrm{insp}$ is chosen uniformly between $0.5$ and $2$. We focus on the late stages of inspiral as these are most likely to be detected by eLISA. Given the waveform, we can estimate the expected SNR by approximating the overlap with a time-domain integral
\begin{equation}
\overlap{s}{h} \approx 4 \sum_{A=+,\times}\intd{0}{T}{\frac{s_A(t)h_A(t)}{S_n(f(t))}}{t},
\end{equation}
where $f(t)$ is some characteristic frequency of the orbit at time $t$. We find that taking $f$ to be twice the (radial) orbital frequency gives a good approximation to the SNR.

To obtain our final distribution of detectable EMRIs, shown as dark blue in \figref{EMRIpar-dists}, we quote the parameter values $t_\mathrm{insp}$ years before plunge, including only those systems with an estimated SNR above 10. Also shown in the figure, in light pink, are the systems with an estimated SNR below 10, to illustrate which systems tend to have lower SNRs. The acceptance rate for generating a detectable EMRI from the entire distribution is $32\%$.

We note that prograde orbits (those with $\cos\iota > 0$) and systems with higher BH spins tend to be favoured; this is likely due to the fact that such systems can inspiral much closer before finally plunging and so experience the strongest gravitational fields. This effect is amplified by the fact that the majority of the SNR is recovered from the final part of the inspiral, which also explains why larger values of $t_\mathrm{insp}$ are only marginally preferred over smaller values.

Systems at higher redshifts start to become undetectable as the amplitude of the signal scales inversely proportional to the luminosity distance to the source. At $z=0.7$, the fraction of detectable systems drops to $5\%$; this is chosen as the upper cutoff in the redshift distribution as a compromise between including as many detectable systems as possible, but without having too low an overall acceptance rate. We quote the masses in the system as "redshifted masses" to account for the FRW background metric.

@@

\begin{enumerate}
\item No significant resonance jumps are observed, and all the adiabatic models agree with the full evolution reasonably well. The overlaps tend to be noticeably lower than $1$ since the adiabatic approach breaks down near the end of the inspiral.

@@

To ensure that our chosen adiabatics are reasonable estimates to maximise the overlap, we computed many adiabatic waveforms on a rigid grid for one particular system; there can be large fluctuations in the overlap, but our chosen parameters perform at least averagely well.

@@

While resonance jumps may in principle be as large as a few percent (see the example in \figref{resjump-vs-q}), this tends to only occur in high eccentricity systems. The typical size of any resonance jump within our astrophysical population is less than $0.01\%$, with many systems showing no visible changes through resonance. \Figref{pop-SNR-vs-jump} shows the distribution of computed resonance jumps in $E$ for our population, along with the corresponding loss of overlap. There is no observed trend with the size of the resonance jump, although this is not surprising given the small size of the jumps. The large scatter on the plot is due to our choice of adiabatic family; in each case, the quoted loss of SNR is a lower bound for the actual loss of SNR across the entire adiabatic model space.

@@

It is possible for EMRIs to form through alternative channels, which may result in a population of high eccentricity systems. For example, 'plunging' compact objects can be mistaken for slow inspirals since they still undergo many hundreds of thousands of cycles in the bandwidth of a detector like eLISA \cite{Amaro-Seoane2013}. This type of system is expected to be orders of magnitude more common than classic EMRIs. It is expected that resonance behaviours would be much more important in this case.

@@

\begin{figure}[htbp]
\centering
\includegraphics[width=0.46\textwidth]{Fig_SNRcutoff}
\caption{\label{fig:SNRcutoff}The fraction of EMRI systems that would be lost using adiabatic models to analyse data that contain the effects of transient resonances. The error bars are estimated using the beta distribution.}
\end{figure}

>>>>

To investigate how resonances would impact real-world observations of EMRIs using GWs, we need to generate a realistic astrophysical population of EMRIs and then evolve these systems in time, including a mechanism to generate the GWs at infinity, which can then be analysed.

@@

\begin{figure}[htbp]
\centering
\includegraphics[width=0.46\textwidth]{Fig_overlap_vs_time}
\caption{\label{fig:overlap-dephasing}The overlap computed between the full evolution and an adiabatic evolution for an illustrative EMRI system with $p=7.85$, as a function of time, including only the parts of the waveforms within some $\Delta t$ of $t$, arbitrarily chosen to give 25 independent (non-overlapping) calculations. The time axis is centred on the value at which the full evolution crosses the 2:3 resonance.}
\end{figure}

@@

To be able to quantify the effect that the resonance has on an evolution, we turn to the SNR and the overlap between full and adiabatic models computed for the entire length of the observed inspiral. The dephasing means that an adiabatic evolution with the same initial conditions as a full evolution is likely to have a low overlap if a significant resonance is encountered. However, this does not necessarily mean that all adiabatic evolutions have low overlaps; we need to search for optimal adiabatic parameters, such that the overlap is maximal.

@@

The most restrictive aspect of this approach is that all extrinsic parameters and any intrinsic ones that remain constant throughout the evolution, such as the masses of the BHs and the central BH spin, do not get explored by our family of adiabatic models. It is possible that small deviations from these true parameters give rise to better matching adiabatic models, to partially correct for the (lack of) resonance effects. Indeed, systematic biases are expected when the model waveforms do not accurately reflect the data~\cite{Cutler2007}.

The problem of searching over adiabatic templates now reduces to the task of choosing appropriate values of $t_{\mathrm{match}}$. Anticipating that resonances shall play a key role in determining the degree of waveform overlap, we choose matching times situated at $5\tau_\mathrm{res}$ after each resonance of interest, namely the low-order 2:3, 1:2, 2:5 and 1:3. These matching times lie in a portion of the evolution that is not affected by a resonance and so should result in a large overlap for that region of the inspiral. We also include templates that match at the start of the evolution, so that we have a model that should match the portion of the evolution before the first resonance, and at the end of the evolution, as this is often where most SNR is accumulated. In addition, we include models that match exactly on each of the resonances as these may achieve partial matching both before and after the resonance and hence perform well.

@@

Given our computed resonance jumps for all of the orbital parameters, we extract a single phase parameter $q$ for each system that specifies the magnitude of each jump. We expect such a phase parameter to be linearly offset from $\widehat{\kappa}_0$. The resonance jumps are sinusoidal with respect to $q$, as shown by the fits in \figref{resjump-vs-q}, lending support to \eqnref{delta-I-a} and giving us confidence in our numerical jump calculations. We can also compute the jumps in the fundamental frequencies for each system (shown as dashed lines in \figref{resjump-vs-q}).

@@

Strong resonances can severely limit our ability to recover SNR from waveforms using adiabatic templates; they partition the inspiral, splitting up the total SNR into different adiabatic modes, which may be individually undetectable. To assess the impact of this on future GW missions, we need to create a sample population of detectable EMRIs.

We first identify 14 parameter distributions that describe an astrophysically typical EMRI\footnote{The actual distribution in parameter space of observable EMRI systems is unknown, but we can make reasonable estimates.}. We choose the mass of the compact object $\mu = 10 M_\odot$, corresponding to a typical mass for stellar BHs. We take the central object to be a typical supermassive BH at the centre of a galaxy, with mass $M$ uniformly distributed in log-space between $10^5 M_\odot$ and $10^7 M_\odot$, and dimensionless spin $a_\ast$ uniformly between its limiting values, $0$ and $1$. The direction of the spin of the central BH, the direction of the angular momentum of the orbit and the direction of the source on the sky are all chosen uniformly on the sphere. The angles specifying the location of pericenter and the compact object around the orbit are distributed uniformly between $0$ and $2\pi$.

@@

These EMRIs are astrophysically motivated, but may not be detectable by eLISA. To assess this, we evolve the systems forward in time using the analytic kludge method of Barack and Cutler \cite{Barack2004} until they reach the last stable orbit. We then calculate the Peters-Matthews waveform \cite{Peters1963} for a time $t_\mathrm{insp}$ at the end of the inspiral, where $t_\mathrm{insp}$ is chosen uniformly between $0.5$ years and $2$ years. We focus on the late stages of inspiral as these are most likely to be detected by eLISA. Given the waveform, we can estimate the expected SNR by approximating the overlap with a time-domain integral
\begin{equation}
\overlap{s}{h} \approx 4 \intd{0}{t_\mathrm{insp}}{\frac{s(t)h(t)}{S_n(f(t))}}{t},
\end{equation}
where $f(t)$ is some characteristic frequency of the orbit at time $t$. We find that taking $f$ to be twice the (radial) orbital frequency gives a good approximation to the SNR.

To obtain our final distribution of detectable EMRIs, shown as dark blue in \figref{EMRIpar-dists}, we quote the parameter values $t_\mathrm{insp}$ before plunge. Only systems with $\rho > 10$ are treated as detectable. Also shown in the figure, in light pink, are the systems with an estimated SNR below 10, to illustrate which systems tend to have lower SNRs. The acceptance rate for generating a detectable EMRI from the entire distribution is $32\%$.

We note that prograde orbits (those with $\cos\iota > 0$) and systems with higher BH spins tend to be favoured; this is likely due to the fact that such systems can inspiral much closer before finally plunging, and so experience the strongest gravitational fields. This effect is amplified by the fact that the majority of the SNR is recovered from the final part of the inspiral, which also explains why larger values of $t_\mathrm{insp}$ are only marginally preferred over smaller values.

Systems at higher redshifts start to become undetectable as the amplitude of the signal scales inversely proportional to the luminosity distance to the source. At $z=0.7$, the fraction of detectable systems drops to $5\%$; this is chosen as the upper limit in the redshift distribution to minimise unneccessary computations. We quote the masses in the system as redshifted masses, which accounts for the frequency shift due to the expansion of the Universe.

@@

\begin{enumerate}
\item No significant resonance jumps are observed, and all the adiabatic models agree with the full evolution reasonably well: the overlaps are typically larger than $0.9$. The models have overlaps less than $1$ since the adiabatic approach breaks down near the end of the inspiral.

@@

To ensure that our chosen adiabatics are reasonable estimates to maximise the overlap, we computed many adiabatic waveforms with regular values of $t_\mathrm{match}$ for one particular system; there can be large fluctuations in the overlap, but our chosen parameters perform at least averagely well.

@@

While resonance jumps may in principle be as large as a few percent (see the example in \figref{resjump-vs-q}), this tends to only occur in high eccentricity systems. The typical size of any resonance jump within our astrophysical population is less than $0.01\%$, with many systems showing no visible changes through resonance. \Figref{pop-SNR-vs-jump} shows the distribution of computed resonance jumps in $E$ for our population, along with the corresponding loss of overlap. There is no observed trend with the size of the resonance jump, although this is not surprising given the small size of the jumps. The large scatter on the plot is due to our choice of adiabatic family; in each case, the quoted fractional SNR is a lower bound for the actual fractional SNR across the entire adiabatic model space.

@@

It is possible for EMRIs to form through alternative channels, which may result in a population of high eccentricity systems. As an example, compact objects can be scattered onto orbits that plunge directly into the central BH; these systems can be mistaken for slow inspirals since they still undergo many hundreds of thousands of cycles in the bandwidth of a detector like eLISA \cite{Amaro-Seoane2013}. This kind of event is expected to be orders of magnitude more common than classic EMRIs. It is expected that resonance behaviours would be much more important in this case.

@@

\begin{figure}[htbp]
\centering
\includegraphics[width=0.46\textwidth]{Fig_SNRcutoff}
\caption{\label{fig:SNRcutoff}The fraction of EMRI systems that would be lost using adiabatic models to analyse data that contain the effects of transient resonances. The error bars are estimated using the beta distribution \cite{Cameron2011}.}
\end{figure}
